### 2016-11-25 simulations/simulation.Rmd

---
title: "ANOVA Comparison Simulations"
author: "Kellie Ottoboni"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE,
                      warning = FALSE,
                      tidy = TRUE, tidy.opts = list(width.cutoff = 60),
                      cache = TRUE)
#devtools::install_github("statlab/permuter")
require(permuter)
require(dplyr)
require(reshape2)
require(ggplot2)
require(xtable)
options(xtable.comment = FALSE, xtable.include.rownames = FALSE, xtable.align = "p{3.5cm}")
require(grid)
require(gridExtra)
```

# The model

We assume an additive linear model:

$$Y_{ij1} =\beta_0Y_{i0} + \beta_{j} + \gamma_j Z_{ij} + \varepsilon_{ij}$$

for individuals $i = 1, \dots, n_j$, $j = 1, \dots, J$.
$\beta_0$ is the coefficient for the baseline measurement $Y_{i0}$, $\beta_j$ is the mean effect of being at site $j$, $Z_{ij}$ is the treatment level, $\gamma_j$ is the effect of treatment at site $j$, and $\varepsilon_{ij}$ is an error term.
We will assume that $\beta_0 = 1$.

Suppose there are three sites with $\beta_1 = 1, \beta_2 = 1.5,$ and $\beta_3 = 2$.
Assume that there are 16 individuals per site and treatment assignment is balanced, i.e. 8 people receive each treatment at each site.

set up 1: $\gamma_1 = \gamma_2 = \gamma_3$
set up 2: $\gamma_1 = \gamma > 0$, $\gamma_2 = \gamma_3 = 0$
errors 1: $\varepsilon \sim N(0, \sigma^2)$
errors 2: $\varepsilon \sim t(2)$

```{r generate_sample}
generate_simulated_data <- function(gamma, effect, errors, n = c(16, 16, 16)){
  # Input:
  # gamma = the magnitude of the treatment effect
  # effect = "same effect" or "single site effect" - which sites have a tr effect > 0?
  # errors = "normal" or "heavy"
  # Returns: a dataframe containing columns named Y1 (response), Y0 (baseline), Z (treatment), and SITEID (stratum)

  SITEID <- rep(1:3, times = n)
  N <- sum(n)
  beta <- c(1, 1.5, 2)

  # What is the treatment effect?
  if(effect == "same effect"){
    gamma_vec <- rep(gamma, N)
  } else {
    gamma_vec <- rep(c(gamma, 0, 0), times = n)
  }

  # Generate errors
  if(errors == "normal"){
    epsilon <- rnorm(N)
  } else {
    epsilon <- rt(N, df = 2)
  }

  # Generate covariates
  Y0 <- rnorm(N)
  Z <- rep(0:1, length.out = N)
  site_effect <- rep(beta, times = n)
  Y1 <- Y0 + gamma_vec*Z + site_effect + epsilon
  return(data.frame(Y1, Y0, Z, SITEID))
}

generate_simulated_pvalues <- function(dataset, reps = 1e3){
  # Inputs:
  # dataset = a dataframe containing columns named Y1 (response), Y0 (baseline), Z (treatment), and SITEID (stratum)
  # Returns: a vector of p-values
  # first element is the p-value from the ANOVA
  # second element is the p-value from the stratified two-sample permutation test
  # third element is the p-value from the linear model test, permuting treatment
  # fourth element is the p-value from the Freedman-Lane linear model test, permuting residuals

  # ANOVA
  modelfit <- lm(Y1 ~ Y0 + Z + factor(SITEID), data = dataset)
  resanova <- summary(aov(modelfit))
  anova_pvalue <- resanova[[1]]["Z", "Pr(>F)"]

  # Permutation test
  observed_diff_means <- mean(dataset$Y1[dataset$Z == 1]) - mean(dataset$Y1[dataset$Z == 0])
  diff_means_distr <- stratified_two_sample(group = dataset$Z, response = dataset$Y1, stratum = dataset$SITEID, reps = reps)
  # diff_means_distr2 <-  replicate(reps, {
  #   Z_perm <- permute_within_groups(dataset$Z, dataset$SITEID)
  #   mean(dataset$Y1[Z_perm == 1]) - mean(dataset$Y1[Z_perm == 0])
  # })
  perm_pvalue <- t2p(observed_diff_means, diff_means_distr, alternative = "two-sided")

  # Diffed permutation test
  dataset$diff <- dataset$Y1 - dataset$Y0
  observed_diff_means2 <- mean(dataset$diff[dataset$Z == 1]) - mean(dataset$diff[dataset$Z == 0])
  diff_means_distr2 <- stratified_two_sample(group = dataset$Z, response = dataset$diff, stratum = dataset$SITEID, reps = reps)
  perm_pvalue2 <- t2p(observed_diff_means2, diff_means_distr2, alternative = "two-sided")

  # Permutation of treatment in linear model
  observed_t1 <- summary(modelfit)[["coefficients"]]["Z", "t value"]
  lm1_t_distr <-  replicate(reps, {
    dataset$Z_perm <- permute_within_groups(dataset$Z, dataset$SITEID)
    lm1_perm <- lm(Y1 ~ Y0 + Z_perm + factor(SITEID), data = dataset)
    summary(lm1_perm)[["coefficients"]]["Z_perm", "t value"]
  })
  lm_pvalue <- t2p(observed_t1, lm1_t_distr, alternative = "two-sided")

  # Freedman-Lane linear model residual permutation
  lm2_no_tr <- lm(Y1~Y0 + factor(SITEID), data = dataset)
  lm2_resid <- residuals(lm2_no_tr)
  lm2_yhat <- fitted(lm2_no_tr)
  lm2_t_distr <-  replicate(reps, {
    lm2_resid_perm <- permute_within_groups(lm2_resid, dataset$SITEID)
    dataset$response_fl <-  lm2_yhat + lm2_resid_perm
    lm2_perm <- lm(response_fl ~ Y0 + Z + factor(SITEID), data = dataset)
    summary(lm2_perm)[["coefficients"]]["Z", "t value"]
  })
  fl_pvalue <- t2p(observed_t1, lm2_t_distr, alternative = "two-sided")

  return(c("ANOVA" = anova_pvalue, 
           "Stratified Permutation" = perm_pvalue, 
           "Diffed Stratified Permutation" = perm_pvalue2,
           "LM Permutation" = lm_pvalue, 
           "Freedman-Lane" = fl_pvalue))
}
```

```{r function_power_curves}
compute_power <- function(pvalues){
  sapply((0:99)/100, function(p) mean(pvalues <= p, na.rm = TRUE))
}

plot_power_curves <- function(power_mat, title){
  melt(power_mat) %>% 
  mutate("pvalue" = Var1/100) %>%
  mutate("Method" = Var2) %>%
  ggplot(aes_string(x = "pvalue", y = "value", color = "Method")) +
  geom_line() +
  xlab("P-value") +
  ylab("Power") +
  ggtitle(title)
}

plot_pvalue_hist <- function(pvalue_mat, title){
  melt(pvalue_mat) %>%
  mutate("Method" = Var2) %>%
  ggplot(aes(x = value, fill = Method)) +
  geom_histogram() +
  facet_wrap(~Method) +
  ggtitle(title)
}

plot_pvalue_scatter <- function(pvalue_mat, title){
  pvalue_mat %>%
  as.data.frame() %>%
  select(ANOVA, strat = starts_with("Stratified")) %>%
  ggplot(aes(x = ANOVA, y = strat)) + 
  geom_point() +
  xlim(0, 1) + ylim(0, 1) +
  ylab("Stratified Permutation") +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
  ggtitle(title)
}
```

```{r design_null, fig.height = 4}
set.seed(760682460) # Generated from random.org Timestamp: 2016-11-14 10:21:12 UTC

design0_pvalues <- replicate(100, {
  tmp <- generate_simulated_data(gamma = 0, effect = "same effect", errors = "normal")
  generate_simulated_pvalues(tmp)
})
design0_pvalues <- t(design0_pvalues)
colnames(design0_pvalues) <- c("ANOVA", "Stratified Permutation", "Diffed Permutation", "LM Permutation", "Freedman-Lane")
design0_power <- apply(design0_pvalues, 2, compute_power)

plot_power_curves(design0_power, "No Effect, Normal Errors")

plot_pvalue_hist(design0_pvalues, "No Effect, Normal Errors")

plot_pvalue_scatter(design0_pvalues, "Simulation P-values for Data with \n No Effect, Normal Errors")
```

```{r design1, fig.height = 4}
set.seed(760682460) # Generated from random.org Timestamp: 2016-11-14 10:21:12 UTC

design1_pvalues <- replicate(100, {
  tmp <- generate_simulated_data(gamma = 1, effect = "same effect", errors = "normal")
  generate_simulated_pvalues(tmp)
})
design1_pvalues <- t(design1_pvalues)
colnames(design1_pvalues) <- c("ANOVA", "Stratified Permutation", "Diffed Permutation", "LM Permutation", "Freedman-Lane")
design1_power <- apply(design1_pvalues, 2, compute_power)

plot_power_curves(design1_power, "Constant Additive Effect of 1, Normal Errors")

plot_pvalue_hist(design1_pvalues, "Constant Additive Effect of 1, Normal Errors")

#plot_pvalue_scatter(design1_pvalues, "Simulation P-values for Data with \n Constant Additive Effect of 1, Normal Errors")
```

```{r design2, fig.height = 4}
set.seed(760682460) # Generated from random.org Timestamp: 2016-11-14 10:21:12 UTC

design2_pvalues <- replicate(100, {
  tmp <- generate_simulated_data(gamma = 1, effect = "single site effect", errors = "normal")
  generate_simulated_pvalues(tmp)
})
design2_pvalues <- t(design2_pvalues)
colnames(design2_pvalues) <- c("ANOVA", "Stratified Permutation", "Diffed Permutation", "LM Permutation", "Freedman-Lane")
design2_power <- apply(design2_pvalues, 2, compute_power)

plot_power_curves(design2_power, "Effect of 1 at a Single Site, Normal Errors")

plot_pvalue_hist(design2_pvalues, "Effect of 1, Normal Errors")

plot_pvalue_scatter(design2_pvalues, "Simulation P-values for Data with \n Effect of 1 at a Single Site, Normal Errors")
```

```{r design3, fig.height = 4}
set.seed(760682460) # Generated from random.org Timestamp: 2016-11-14 10:21:12 UTC

design3_pvalues <- replicate(100, {
  tmp <- generate_simulated_data(gamma = 1, effect = "same effect", errors = "heavy")
  generate_simulated_pvalues(tmp)
})
design3_pvalues <- t(design3_pvalues)
colnames(design3_pvalues) <- c("ANOVA", "Stratified Permutation", "Diffed Permutation", "LM Permutation", "Freedman-Lane")
design3_power <- apply(design3_pvalues, 2, compute_power)

plot_power_curves(design3_power, "Constant Additive Effect of 1, Heavy-tailed Errors")

plot_pvalue_hist(design3_pvalues, "Constant Additive Effect of 1, Heavy-tailed Errors")

plot_pvalue_scatter(design3_pvalues, "Simulation P-values for Data with \n Constant Additive Effect of 1, Heavy-tailed Errors")
```

```{r design4, fig.height = 4}
set.seed(760682460) # Generated from random.org Timestamp: 2016-11-14 10:21:12 UTC

design4_pvalues <- replicate(100, {
  tmp <- generate_simulated_data(gamma = 1, effect = "single site effect", errors = "heavy")
  generate_simulated_pvalues(tmp)
})
design4_pvalues <- t(design4_pvalues)
colnames(design4_pvalues) <- c("ANOVA", "Stratified Permutation", "Diffed Permutation", "LM Permutation", "Freedman-Lane")
design4_power <- apply(design4_pvalues, 2, compute_power)

plot_power_curves(design4_power, "Effect of 1 at a Single Site, Heavy-tailed Errors")

plot_pvalue_hist(design4_pvalues, "Effect of 1, Heavy-tailed Errors")

plot_pvalue_scatter(design4_pvalues, "Simulation P-values for Data with \n Effect of 1 at a Single Site, Heavy-tailed Errors")
```