### 2016-12-08 ms/fig/normal_simulation_power_summary.tex
Rerun sims with 1000 reps to get cleaner plots and power estimates

\begin{table}[ht]
\centering
\begin{tabular}{p{1.15in}|p{0.7in}|p{0.6in}p{0.8in}p{0.8in}p{0.8in}p{0.75in}}
  \hline
Treatment & Errors & ANOVA & Stratified Permutation & Differenced Permutation & LM Permutation & Freedman-Lane \\ 
  \hline
Constant Additive Effect & Gaussian & 0.89 & 0.55 & 0.86 & 0.89 & 0.89 \\ 
  Single Stratum Effect & Gaussian & 0.16 & 0.09 & 0.16 & 0.16 & 0.17 \\ 
  Constant Additive Effect & Heavy-tailed & 0.35 & 0.24 & 0.38 & 0.40 & 0.39 \\ 
  Single Stratum Effect & Heavy-tailed & 0.07 & 0.08 & 0.08 & 0.08 & 0.08 \\ 
   \hline
\end{tabular}
\caption{Empirical power at level $0.05$ for Gaussian simulated data} 
\label{tab:normal_power}
\end{table}