### 2016-12-08 ms/fig/skewed_simulation_power_summary.tex

\begin{table}[ht]
\centering
\begin{tabular}{p{1.25in}|p{0.7in}|p{0.6in}p{0.8in}p{0.8in}p{0.8in}p{0.75in}}
  \hline
Treatment & Design & ANOVA & Stratified Permutation & Differenced Permutation & LM Permutation & Freedman-Lane \\ 
  \hline
Constant Additive Effect & Balanced & 0.97 & 0.14 & 0.85 & 0.97 & 0.98 \\ 
  Single Stratum Effect & Balanced & 0.42 & 0.05 & 0.28 & 0.43 & 0.42 \\ 
  Constant Additive Effect & Imbalanced & 0.97 & 0.15 & 0.86 & 0.97 & 0.97 \\ 
  Single Stratum Effect & Imbalanced & 0.19 & 0.07 & 0.10 & 0.19 & 0.19 \\ 
   \hline
\end{tabular}
\caption{Empirical power at level $0.05$ for discrete, skewed simulated data} 
\label{tab:skewed_power}
\end{table}