### 2016-12-19 ms/fig/imbalanced_simulation_power_summary.tex
Add simulations for imbalanced design; update tables and figs; edits to results

\begin{table}[ht]
\centering
\begin{tabular}{|p{0.7in}|p{0.6in}p{0.8in}p{0.8in}p{0.8in}p{0.8in}p{0.75in}}
  \hline
Treatment & Errors & ANCOVA & Stratified Permutation & Differenced Permutation & LM Permutation & Freedman-Lane \\ 
  \hline
No Effect & Gaussian & 0.046 & 0.031 & 0.028 & 0.042 & 0.045 \\ 
  Constant Additive Effect & Gaussian & 0.840 & 0.490 & 0.810 & 0.870 & 0.860 \\ 
  Single Stratum Effect & Gaussian & 0.120 & 0.050 & 0.090 & 0.100 & 0.120 \\ 
   \hline
No Effect & Heavy-tailed & 0.034 & 0.054 & 0.055 & 0.047 & 0.042 \\ 
  Constant Additive Effect & Heavy-tailed & 0.230 & 0.200 & 0.270 & 0.260 & 0.270 \\ 
  Single Stratum Effect & Heavy-tailed & 0.070 & 0.080 & 0.090 & 0.090 & 0.080 \\ 
   \hline
\end{tabular}
\caption{Empirical power at level $0.05$ for Gaussian simulated data with imbalanced treatment groups} 
\label{tab:imbalanced_power}
\end{table}