### 2016-12-19 ms/fig/skewed_simulation_power_summary.tex

\begin{table}[ht]
\centering
\begin{tabular}{p{1.25in}|p{0.7in}|p{0.6in}p{0.8in}p{0.8in}p{0.8in}p{0.75in}}
  \hline
Treatment & Design & ANCOVA & Stratified Permutation & Differenced Permutation & LM Permutation & Freedman-Lane \\ 
  \hline
No Effect & Balanced & 0.036 & 0.025 & 0.021 & 0.035 & 0.037 \\ 
  Constant Additive Effect & Balanced & 0.974 & 0.143 & 0.853 & 0.972 & 0.977 \\ 
  Single Stratum Effect & Balanced & 0.422 & 0.050 & 0.275 & 0.429 & 0.424 \\ 
   \hline
No Effect & Imbalanced & 0.038 & 0.024 & 0.019 & 0.037 & 0.039 \\ 
  Constant Additive Effect & Imbalanced & 0.971 & 0.147 & 0.860 & 0.970 & 0.972 \\ 
  Single Stratum Effect & Imbalanced & 0.190 & 0.070 & 0.100 & 0.190 & 0.190 \\ 
   \hline
\end{tabular}
\caption{Empirical power at level $0.05$ for discrete, skewed simulated data} 
\label{tab:skewed_power}
\end{table}