### 2017-01-31 ms/fig/normal_simulation_power_summary.tex
Reformat tables nicely

\begin{table}[ht]
\centering
\begin{tabular}{p{1.15in}|p{0.7in}|p{0.6in}p{0.8in}p{0.8in}p{0.8in}p{0.75in}}
  \hline
Errors & Treatment & ANCOVA & Stratified Permutation & Differenced Permutation & LM Permutation & Freedman-Lane \\ 
  \hline
Gaussian & No Effect & 0.058 & 0.042 & 0.038 & 0.062 & 0.057 \\ 
  Gaussian & Constant Additive Effect & 0.886 & 0.545 & 0.858 & 0.891 & 0.892 \\ 
  Gaussian & Single Stratum Effect & 0.159 & 0.089 & 0.160 & 0.162 & 0.170 \\ 
   \hline
Heavy-tailed & No Effect & 0.032 & 0.040 & 0.045 & 0.042 & 0.040 \\ 
  Heavy-tailed & Constant Additive Effect & 0.355 & 0.240 & 0.378 & 0.400 & 0.393 \\ 
  Heavy-tailed & Single Stratum Effect & 0.066 & 0.080 & 0.082 & 0.076 & 0.080 \\ 
   \hline
\end{tabular}
\caption{Empirical power at level $0.05$ for Gaussian simulated data} 
\label{tab:normal_power}
\end{table}