### 2017-09-05 simulations/continuous_outcomes.Rmd
Create new simulations with continuous outcomes in potential outcome framework

---
title: "ANCOVA Comparison Simulations: Continuous Outcomes"
author: "Kellie Ottoboni"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE,
                      warning = FALSE,
                      tidy = TRUE, tidy.opts = list(width.cutoff = 60),
                      cache = TRUE)
#devtools::install_github("statlab/permuter")
require(permuter)
require(dplyr)
require(reshape2)
require(ggplot2)
require(xtable)
options(xtable.comment = FALSE, xtable.include.rownames = FALSE, xtable.align = "p{3.5cm}")
require(grid)
require(gridExtra)

report_theme <- theme_bw() + theme(
  axis.text.x = element_text(size = 14, angle = -45),
  axis.text.y = element_text(size = 14),
  axis.title = element_text(size = 20),
  title = element_text(size = 20),
  legend.title = element_text(size = 14),
  legend.text = element_text(size = 14),
  strip.text.x = element_text(size = 14)
)
```

# Continuous data

Step 1:
We generated continuous potential outcomes.
We used two scenarios.
In the first, the treatment effect was constant across strata.
We drew a latent random variate $v_{ij}$ from the uniform distribution on $[-4, 4]$.
In the second scenario, the treatment effect varied across strata, and we drew the latent random variable according to
\begin{displaymath}
   v_{ij} \sim \left\{
     \begin{array}{lr}
       \text{Unif}[-4, -1] & : j=1\\
       \text{Unif}[-1, 1] & : j=2\\
       \text{Unif}[1, 4] & : j=3\\
     \end{array}
   \right.
\end{displaymath}

We generated independent and identically distributed errors $\varepsilon_{ij}$ and $\delta_{ij}$ and varied the error distribution. 
The errors were either standard normal (to mimic the usual ANCOVA assumptions),
$t$ distributed with two degrees of freedom,
standard lognormal,
or exponentially distributed, with scale parameter $1$ and shifted to have mean zero.
The observed $(v_{ij}, \varepsilon_{ij}, \delta_{ij})$ were independent across $i$ and $j$.

The baseline value for individual $i,j$ was 
$$ X_{ij} = \frac{\gamma e^{v_{ij}} + e^{v_{ij}/2}}{2} + \varepsilon_{ij}$$.

Then we generated potential outcomes as

$$Y_{ij}(Z_{ij}) = \frac{(2Z_{ij} - 1)\gamma e^{v_{ij}} + e^{v_{ij}/2}}{2} + \delta_{ij}.$$

The treatment effect for individual $(i, j)$ is $\gamma e^{v_{ij}}$ and we would like to estimate the average treatment effect in the sample.

Assume that there are $n_j=16$ individuals per stratum and treatment assignment is balanced, i.e. 8 people receive each treatment at each stratum.
After sampling $(v_{ij}, \varepsilon_{ij}, \delta_{ij})$, we regenerate $Z$ 10,000 times.
We repeat this procedure for each distribution of latent variables $v$ and of the errors $\varepsilon$ and $\delta$.

Step 2:
We regenerate $Z$ and recompute $Y_(Z)$ 10,000 times for each design.

Step 3:
We repeat step 1 and 2 for different distributions on $\varepsilon$ and $\delta$.

In expectation, the average treatment effect is $\gamma$.
We compare the empirical power of five tests to detect this treatment effect:

* ANCOVA: we fit a linear model of response $Y_1$ on baseline $Y_0$, treatment $Z$, and a dummy for stratum.
* Stratified permutation: we permute treatment assignment within stratum, then take the difference in means between treated and control outcomes $Y_1$
* Differenced permutation: we do the same permutation procedure as the stratified permutation test, except we use the difference between outcome and baseline, $Y_1 - Y_0$
* Linear model (LM) permutation: we use the same stratified permutation procedure as above, except use the $t$-statistic for the coefficient on treatment in the linear regression of $Y_1$ on $Y_0$, $Z$, and stratum dummies
* Freedman-Lane test: see the other Rmd document for a full description of this procedure

# Data-generation, tests, and plotting functions

```{r generate_sample}
gen_y1 <- function(gamma, v, error, Z){
  y1 <- 0.5*((2*Z-1)*gamma*exp(v) + exp(v/2)) + error
  return(y1)
}

gen_y0 <- function(gamma, v, error){
  y0 <- 0.5*(-1*gamma*exp(v) + exp(v/2)) + error
  return(y0)
}

generate_simulated_data <- function(gamma, effect, errors, n = c(16, 16, 16)){
  # Input:
  # gamma = multiplier for the magnitude of the treatment effect
  # effect = "same effect" or "heterogeneous"
  # errors = "normal" or "heavy"
  # n = number of individuals at each stratum
  # Returns: a dataframe containing columns named Y1 (response), Y0 (baseline), Z (treatment), gamma_vec (treatment effect per individual), stratumID (stratum), stratum_effect (beta coefficient per individual), and epsilon (errors)

  stratumID <- rep(1:3, times = n)
  N <- sum(n)

  # What is the treatment effect?
  if(effect == "same effect"){
    v <- runif(N, min=-4, max=4)
  } else if(effect == "heterogeneous"){
    v <- rep(0, N)
    v[stratumID == 1] <- runif(n[1], min=-4, max=-1)
    v[stratumID == 2] <- runif(n[2], min=-1, max=1)
    v[stratumID == 3] <- runif(n[3], min=1, max=4)
  } else {
    stop("invalid parameter effect")
  }

  # Generate errors
  if(errors == "normal"){
    epsilon <- rnorm(N)
    delta <- rnorm(N)
  } else if (errors == "t"){
    epsilon <- rt(N, df = 2)
    delta <- rt(N, df = 2)
  } else if (errors == "lognormal"){
    epsilon <- rlnorm(N)
    delta <- rlnorm(N)
  } else if (errors == "exponential"){
    epsilon <- rexp(N) - 1
    delta <- rexp(N) - 1
  } else {
    stop("invalid errors parameter")
  }

  # Generate covariates
  Z <- rep(0:1, length.out = N)
  Y0 <- gen_y0(gamma, v, epsilon)
  Y1 <- gen_y1(gamma, v, delta, Z)
  return(data.frame(Y1, Y0, Z, v, stratumID, epsilon, delta))
}

generate_simulated_pvalues <- function(dataset, reps = 1e3){
  # Inputs:
  # dataset = a dataframe containing columns named Y1 (response), Y0 (baseline), Z (treatment), and stratumID (stratum)
  # Returns: a vector of p-values
  # first element is the p-value from the ANCOVA
  # second element is the p-value from the stratified two-sample permutation test
  # third element is the p-value from the linear model test, permuting treatment
  # fourth element is the p-value from the Freedman-Lane linear model test, permuting residuals

  # ANCOVA
  modelfit <- lm(Y1 ~ Y0 + Z + factor(stratumID), data = dataset)
  resanova <- summary(aov(modelfit))
  anova_pvalue <- resanova[[1]]["Z", "Pr(>F)"]

  # Stratified permutation test of Y1
  observed_diff_means <- mean(dataset$Y1[dataset$Z == 1]) - mean(dataset$Y1[dataset$Z == 0])
  diff_means_distr <- stratified_two_sample(group = dataset$Z, response = dataset$Y1, stratum = dataset$stratumID, reps = reps)
  perm_pvalue <- t2p(observed_diff_means, diff_means_distr, alternative = "two-sided")

  # Diffed permutation test of Y1-Y0
  dataset$diff <- dataset$Y1 - dataset$Y0
  observed_diff_means2 <- mean(dataset$diff[dataset$Z == 1]) - mean(dataset$diff[dataset$Z == 0])
  diff_means_distr2 <- stratified_two_sample(group = dataset$Z, response = dataset$diff, stratum = dataset$stratumID, reps = reps)
  perm_pvalue2 <- t2p(observed_diff_means2, diff_means_distr2, alternative = "two-sided")

  # Permutation of treatment in linear model
  observed_t1 <- summary(modelfit)[["coefficients"]]["Z", "t value"]

  # Freedman-Lane linear model residual permutation
  lm2_no_tr <- lm(Y1~Y0 + factor(stratumID), data = dataset)
  dataset$lm2_resid <- residuals(lm2_no_tr)
  lm2_yhat <- fitted(lm2_no_tr)

  lm1and2_t_distr <-  replicate(reps, {
    dataset[, c("Z_perm", "lm2_resid_perm")] <- permute_within_groups(dataset[, c("Z", "lm2_resid")], dataset$stratumID)
    lm1_perm <- lm(Y1 ~ Y0 + Z_perm + factor(stratumID), data = dataset)

    dataset$response_fl <-  lm2_yhat + dataset$lm2_resid_perm
    lm2_perm <- lm(response_fl ~ Y0 + Z + factor(stratumID), data = dataset)

    c(summary(lm1_perm)[["coefficients"]]["Z_perm", "t value"],
      summary(lm2_perm)[["coefficients"]]["Z", "t value"])
  })
  lm_pvalue <- t2p(observed_t1, lm1and2_t_distr[1, ], alternative = "two-sided")
  fl_pvalue <- t2p(observed_t1, lm1and2_t_distr[2, ], alternative = "two-sided")

  return(c("ANCOVA" = anova_pvalue, 
           "Stratified Permutation" = perm_pvalue, 
           "Differenced Permutation" = perm_pvalue2,
           "LM Permutation" = lm_pvalue, 
           "Freedman-Lane" = fl_pvalue))
}
```

```{r function_power_curves}
compute_power <- function(pvalues){
  sapply((0:99)/100, function(p) mean(pvalues <= p, na.rm = TRUE))
}

plot_power_curves <- function(power_mat, title){
  melt(power_mat) %>% 
  mutate("pvalue" = Var1/100) %>%
  mutate("Method" = Var2) %>%
  ggplot(aes_string(x = "pvalue", y = "value", color = "Method")) +
  geom_line() +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
  xlab("P-value") +
  ylab("Power") +
  ggtitle(title) +
  theme_bw() +
  theme(
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    axis.title = element_text(size = 16),
    title = element_text(size = 16),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 14),
    strip.text.x = element_text(size = 12)
  ) 
}

plot_power_gamma <- function(powermat_list, gamma_vec, alpha, title){
  gamma_power_alpha <- t(sapply(powermat_list, function(x) x[floor(alpha*nrow(x)),]))
  colnames(gamma_power_alpha) <- c("ANCOVA", "Stratified Permutation", "Differenced Permutation", "LM Permutation", "Freedman-Lane")
  melt(gamma_power_alpha, value.name="Power") %>%
  mutate("Method" = Var2) %>%
  mutate("Effect" = rep(gamma_vec, 5)) %>% 
  ggplot(aes(x = Effect, y = Power)) +
  geom_line(aes(color = Method))+
  xlab("Effect size") +
  theme_bw() +
  theme(
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    axis.title = element_text(size = 16),
    title = element_text(size = 16),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 14),
    strip.text.x = element_text(size = 12)
  ) 
}

```

```{r vary_effect_size_plot, fig.height=4}
load("continuous_outcomes/vary_gamma_results.Rda")
gamma_vec <- seq(0, 0.2, by=0.05)
gamma_power <- lapply(gamma_res, function(x) apply(x, 2, compute_power))
plot_power_gamma(gamma_power, gamma_vec, 0.05, "Power at level 5%, Constant Treatment Effect") 
```

```{r vary_effect_size_het_plot, fig.height=4}
load("continuous_outcomes/vary_gamma_het_results.Rda")
gamma_vec <- seq(0, 0.2, by=0.05)
gamma_power_het <- lapply(gamma_res_het, function(x) apply(x, 2, compute_power))
plot_power_gamma(gamma_power_het, gamma_vec, 0.05, "Power at level 5%, Heterogeneous Treatment Effect") 
```
```{r gaussian_homogeneous}
gaussian_homogeneous <- gamma_power[[which(abs(gamma_vec - 0.2) < 1e-6)]]
```
```{r gaussian_heterogeneous}
load("continuous_outcomes/gaussian_heterogeneous.Rda")
```
```{r t_homogeneous}
load("continuous_outcomes/t_homogeneous.Rda")
```
```{r t_heterogeneous}
load("continuous_outcomes/t_heterogeneous.Rda")

```
```{r ln_homogeneous}
load("continuous_outcomes/lognormal_homogeneous.Rda")

```
```{r ln_heterogeneous}
load("continuous_outcomes/lognormal_heterogeneous.Rda")
```
```{r exponential_homogeneous}
load("continuous_outcomes/exponential_homogeneous.Rda")
```
```{r exponential_heterogeneous}
load("continuous_outcomes/exponential_heterogeneous.Rda")

```

```{r power_summary, results = "asis"}
powers <- list(gaussian_homogeneous,
               t_homogeneous,
               lognormal_homogeneous,
               exponential_homogeneous,
               gaussian_heterogeneous,
               t_heterogeneous,
               lognormal_heterogeneous,
               exponential_heterogeneous)
summary05 <- t(sapply(powers, function(x) x[5,]))
summary05 <- data.frame(rep(c("Normal", "t(2)", "Log Normal", "Exponential"), 2), summary05)
colnames(summary05) <-  c("Errors", "ANCOVA", "Stratified Permutation", "Differenced Permutation", "LM Permutation", "Freedman-Lane") 
```

```{r power_summary_const_tab, results = "asis"}
summarytab_constant <- xtable(summary05[1:4,], digits = 3,
       caption = "Empirical power at level $0.05$ for simulated data with constant additive treatment effects",
       label = "tab:power_grid1") 
print(summarytab_constant, include.rownames=FALSE) 
```

```{r power_summary_het_tab, results = "asis"}
summarytab_het <- xtable(summary05[5:8, ], digits = 3,
       caption = "Empirical power at level $0.05$ for simulated data with heterogeneous treatment effects",
       label = "tab:power_grid2") 
print(summarytab_het, include.rownames=FALSE) 
```